---
title: "prepare input files for H-MAGMA"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
date: "2025-05-14"
---


* the codes are from nature protocol paper [Annotating genetic variants to target genes
using H-MAGMA](https://www.nature.com/articles/s41596-022-00745-z). 



## Prepare libraries and directory ● Timing 5 min


```{r, message=F, warning=F}
rm(list=ls())
set.seed(123)
options(stringsAsFactors=F)
library(GenomicRanges)
library(biomaRt)
library(dplyr)
```




## Create GenomicRanges (GRanges) objects for exon and promoter coordinates ● Timing ~5 min

```{r, message=F, warning=F}
# 3 Read in exonic and promoter coordinate ﬁles by entering the following commands
path="C:\\han\\Dataset\\20250221_PMID_32152537_hmagma\\6382668\\HMAGMA_Protocol_v1.02\\HMAGMA_Protocol\\"
path2="C:\\han\\Projects\\2025_03_breast_cancer\\HMAGMA\\"
path3="C:\\han\\Projects\\2025_03_breast_cancer\\data_for_hmagma\\"
path4="C:\\han\\Dataset\\20251106_PancanQTL_v2\\BRCA.eQTL-GWAS.xls\\"


exon <- read.table(paste0(path, "Gencode26_exon.bed"), header = FALSE)
dim(exon)
head(exon)
```



```{r, message=F, warning=F}

file.exists(paste0(path, "Gencode26_exon.bed")) # check if the file exists 

exon$V1 <- sub("^", "chr", exon$V1)
promoter <- read.table(paste(path, "Gencode26_promoter.bed", sep=""))
promoter$V1 <- sub("^", "chr", promoter$V1)

# 4 Create a GRanges object for exon and promoter deﬁnitions.
exonranges <- GRanges(exon[,1],IRanges(exon[,2],exon[,3]),gene=exon[,4])
promoterranges <- GRanges(promoter[,1], IRanges(promoter[,2], promoter[,3]), gene=promoter[,4])

# 5 Save exon and promoter GRanges objects as a .rda ﬁle for future use.
#save(exonranges, promoterranges, file=paste(path2, "exon_promoranges.rda", sep=""))

exonranges
promoterranges
```


* "Obtain gene and exon coordinates from Gencode version 26 by using the [following link](https://www.
gencodegenes.org/human/release_26lift37.html). We downloaded the GFT ﬁle for basic gene
annotation and converted them to .bed ﬁles. Promoters were deﬁned as 2 kb upstream to every
transcription start site. Gene, exon and promoter coordinate ﬁles are listed as Gencode26_gene.bed,
Gencode26_exon.bed and Gencode26_promoter.bed, respectively, in the Zenodo repository. Columns
for these ﬁles are deﬁned as [chromosome, start, end, gene]." **Genome assembly hg19 was used**. 







## Generate a GRanges object for SNP coordinates ● Timing ~5 min

```{r, message=F, warning=F}
#Generate a GRanges object for SNP coordinates ● Timing ~5 min
#6 Read in the SNP annotation ﬁle by using the following command.
snps <- read.table(paste(path, "EUR.bim.gz", sep=""))
snps <- snps[,c(1,2,4)]
colnames(snps) <- c("chr","SNP","Position")
snps$chr <- sub("^", "chr", snps$chr)


#7 Create a GRanges object for SNP annotations.
snps <- GRanges(snps$chr, IRanges(snps$Position, snps$Position), rsid=
snps$SNP)
#8 Save the SNP GRanges object as a .rda ﬁle for future use.
#save(snps,file=paste(path2, "snps.rda", sep=""))
snps
```


* reference genome, `EUR.bim` is from  European ancestry



## Assign SNPs (in EUR.bim) to genes by overlapping SNPs with exons and promoters ● Timing 10–15 min

```{r, message=F, warning=F}
#9 Identify genes that map to SNPs residing in exons by overlapping the GRanges object for SNPs with
#the GRanges object for exons.
olap <- findOverlaps(snps,exonranges)
snpexon <- snps[queryHits(olap)]
mcols(snpexon) <- cbind(mcols(snpexon), mcols(exonranges[subjectHits(olap)]))
snpexon <- snpexon[seqnames(snpexon)!="chrX"]


#10 Overlap the GRanges object for promoters with the GRanges object for SNPs. Similar to the code
#above, this command will identify genes that map to SNPs residing in promoter regions.
olap <- findOverlaps(snps,promoterranges)
snpro <- snps[queryHits(olap)]
mcols(snpro) <- cbind(mcols(snpro),mcols(promoterranges[subjectHits(olap)]))
snpro <- snpro[seqnames(snpro)!="chrX"]

#11 Save SNPs overlapping with exons and promoters as a .rda ﬁle.
#save(snpro, snpexon, file=paste(path3, "snp_locating_in_exon_promoter_transcript_level.rda", sep=""))
```


* “Find SNPs that fall inside any exon, attach the exon’s annotation (like gene ID) to those SNPs, and remove SNPs on chromosome X.”


## Assign unmapped SNPs (in EUR.bim) to genes on the basis of Hi-C interaction data ● Timing
30–45 min

```{r, message=F, warning=F}
# 12 After identiﬁcation of SNPs that reside in exons and/or promoters, there will be a subset of SNPs
# that do not overlap with either (e.g., intergenic or intronic SNPs). In the following steps, we match
# those SNPs to cognate genes on the basis of Hi-C interaction data. Identify unmapped SNPs by
#using the code below.

snpranges <- snps[!(snps$rsid %in% snpexon$rsid),]
snpranges <- snpranges[!(snpranges$rsid %in% snpro$rsid),]

snpranges
#13 Save unmapped SNPs as a .rda ﬁle.
#save(snpranges, file=paste(path2, "non_exonic_promoter_snp.rda", sep=""))
```



* a majority of SNPs are unmapped (8916368/9822123=90.78%). 


```{r, message=F, warning=F, eval=T}
#14 Read in adult brain Hi-C data. The ﬁrst three columns represent genomic coordinates [chr, start,
#end] of loop anchor 1, and the last three columns represent genomic coordinates of loop anchor 2.
#CRITICAL STEP Although we use a Hi-C dataset from the adult human brain for this protocol,
#Hi-C datasets from the tissue or cell type associated with the disease/trait of interest can be used
#instead.
#hic <- read.table(paste(path, "adultbrain_hic.bedpe", sep=""), header=T)
hic <- read.table(paste0(path3, "PT12345_mapped_5000bp_loops_5000bp_6col.bedpe"))


#15 Restructure Hi-C data to account for both anchors 1 and 2 by using the command below.
hic.int1 <- hic[,1:6]
hic.int2 <- hic[,c(4:6,1:3)]
colnames(hic.int1) = colnames(hic.int2) = c("chrom1", "start1", "end1", "chrom2", "start2", "end2")
hic.comb <- rbind(hic.int1, hic.int2)
hic.comb=hic.comb[-1,]
head(hic.comb)

########### check if  hic.comb has NA or non-numeric values and remove them 
sum(is.na(hic.comb$start1))
sum(is.na(hic.comb$end1))
sum(is.na(hic.comb$start2))
sum(is.na(hic.comb$end2))

bad_rows <- !grepl("^[0-9]+$", hic.comb$start1) | !grepl("^[0-9]+$", hic.comb$end1)
sum(bad_rows)
hic.comb[bad_rows, ]
hic.comb=hic.comb[-which(bad_rows==T),]

#16 Generate a GRanges object for Hi-C data by using the command below.
hicranges <- GRanges(hic.comb$chrom1, IRanges(as.numeric(hic.comb$start1), as.numeric(hic.comb$end1)), int1=hic.comb$start2,int2=hic.comb$end2)
hicranges
```

* the original code only used the first anchor of Hi-C loops, **why don't use the second anchor information or both anchors**  second anchor was used in the next step. 




```{r, message=F, warning=F}
#17 Identify promoter-anchored interactions by overlapping loop anchor 1 with promoters.
olap <- findOverlaps(hicranges,promoterranges)
generanges <- hicranges[queryHits(olap)]
mcols(generanges) <- cbind(mcols(hicranges[queryHits(olap)]), mcols(promoterranges[subjectHits(olap)]))
generanges
```


* **only anchor 1 was used to interact with promoters**



```{r, message=F, warning=F}

#18 Reverse the order of the GRanges object from Step 17. The generanges object created in Step 17
#is in a format of loop anchor 1, loop anchor 2, gene name. We reverse the order of loop
#anchors so that the resulting genebed object has a format of loop anchor 2, loop anchor 1,
#gene name.
genebed <- data.frame(chr=seqnames(generanges), snp.start=generanges$int1, snp.end=generanges$int2, gene.start=start(generanges), gene.end=start(generanges)+width(generanges)-1, ensg=generanges$gene)  # anchor 2 information of int 1,2 are used.  

genebed <- unique(genebed)
str(genebed)
head(genebed)

genebed$snp.start <- as.numeric(as.character(genebed$snp.start))
genebed$snp.end   <- as.numeric(as.character(genebed$snp.end))

#19 Create a GRanges object from Step 18.
genesnpranges <- GRanges(genebed$chr, IRanges(genebed$snp.start,genebed$snp.end), ensg=genebed$ensg)  # anchor 2 information is used 
genesnpranges
```

* `genesnpranges` ranges information is from anchor 2, but ensg from anchor 1  



* **why step 18 is necessary??**




```{r, message=F, warning=F}

#20 Overlap unmapped SNPs from Step 12 with loop anchor 2 from Step 19. This step assigns
#SNPs (located at loop anchor 2) to the genes with which they interact (located at loop
#anchor 1).
olap <- findOverlaps(snpranges,genesnpranges)
snpint <- snpranges[queryHits(olap)]
mcols(snpint) <- cbind(mcols(snpranges[queryHits(olap)]), mcols(genesnpranges[subjectHits(olap)]))
snpint
#save(snpint, file=paste(path3, "PT_Hi-C_transcript_interacting_snp.rda", sep=""))
#file.exists(paste0(path3, "PT_Hi-C_transcript_interacting_snp.rda"))
```




##  ????????? incorporate eQTL and other regulatory information to help map SNPs to genes. 



```{r, message=F, warning=F}
#21 Integrate SNP-gene relationships derived from exons, promoters and Hi-C interaction data.
load(paste(path3, "PT_Hi-C_transcript_interacting_snp.rda", sep=""))
load(paste(path3, "snp_locating_in_exon_promoter_transcript_level.rda", sep=""))
snpdat <- data.frame(chr=seqnames(snpint),cbp=start(snpint), rsid=snpint$rsid, ensg=snpint$ensg)
head(snpdat)

snpromat <- unique(data.frame(rsid=snpro$rsid, ensg=snpro$gene))
snpexonmat <- unique(data.frame(rsid=snpexon$rsid, ensg=snpexon$gene))
snpcomb <- unique(rbind(snpdat[,3:4], snpromat, snpexonmat))
head(snpcomb)
#save(snpcomb, file=paste(path3, "PT12345_Hi-C_SNP_to_transcript_comb.rda", sep=""))



```





```{r, message=F, warning=F, eval=F}
#22 Aggregate SNP-gene relationships to generate the variant-gene annotation ﬁle compatible
#with MAGMA.
snpagg <- aggregate(snpcomb, list(snpcomb$ensg), unique)


#23 Read in the gene deﬁnition ﬁle.
genedef <- read.table(paste(path, "Gencode26_gene.bed", sep=""))
colnames(genedef) <- c("chr", "start", "end", "ensg")


#24 Create an index column from the gene deﬁnition ﬁle consisting of gene chromosomal location, start
#and end.
genedef <- genedef [grep("chr", genedef$chr),]
genedef$chr <- unlist(lapply(strsplit(genedef$chr, "chr"), '[[', 2))
genedef$index <- paste(genedef$chr, genedef$start, genedef$end, sep=":")


#25 Attach the index column from Step 24 to the variant-gene annotation ﬁle from Step 22.
snpagg$index <- genedef[match(snpagg$ensg, genedef$ensg),"index"]


#26 Remove any missing values from the variant-gene annotation ﬁle.
snpagg <- snpagg[!is.na(snpagg$index),]

#27 Subset gene, gene location and SNPs from the variant-gene annotation ﬁle.
snpannot <- snpagg[,c("ensg", "index", "rsid")]

#28 Save the variant-gene annotation ﬁle in an executable format.
writable <- format(snpannot)
write.table(writable, file=paste(path3, "PT12345_SNP_aggregate_transcript.txt", sep=""), quote=F, row.names=F, col.names=F, sep="\t")

#system(paste0("sed -e `s/, /\t/g` < SNP_aggregate_transcript",".txt >","Adultbrain.transcript.annot"))
system("sed -e `s/, /\\t/g` < PT12345_SNP_aggregate_transcript.txt > PT12345_BRSTCancer.transcript.annot")


```


```{r, message=F, warning=F, eval=F}
txt <- readLines(paste0(path3, "PT12345_SNP_aggregate_transcript.txt"))
txt <- gsub(", ", "\t", txt)
writeLines(txt, paste0(path3,"PT12345_BRSTCancer.transcript.annot"))
```


## Check SNP summary data format


### check p values 

```{r, message=F, warning=F}
BRCA_GWAS_eQTL=read.table(paste0(path4, "BRCA.eQTL-GWAS.xls"), header=T, sep = "\t", fill = TRUE, quote = "")

head(BRCA_GWAS_eQTL)
# 1. Is the column numeric?
is.numeric(BRCA_GWAS_eQTL$P_value)

# 2. Are there any NAs, NaNs, or Infs?
sum(is.na(BRCA_GWAS_eQTL$P_value))
sum(is.nan(BRCA_GWAS_eQTL$P_value))
sum(is.infinite(BRCA_GWAS_eQTL$P_value))

# 3. Are all p-values between 0 and 1?
range(BRCA_GWAS_eQTL$P_value, na.rm = TRUE)

sum(BRCA_GWAS_eQTL$P_value>1)
head(BRCA_GWAS_eQTL[which(BRCA_GWAS_eQTL$P_value>1),])

BRCA_GWAS_eQTL2=BRCA_GWAS_eQTL %>% filter(P_value<1) 
BRCA_GWAS_eQTL2$P_value[BRCA_GWAS_eQTL2$P_value==0]=1e-100

sum(is.na(BRCA_GWAS_eQTL2$P_value))



# 1. Select and rename the required columns (SNP ID and P-value)
magma_input <- BRCA_GWAS_eQTL2[c("GWAS_SNP", "P_value")]

# 2. **CRITICAL STEP: Rename the columns to MAGMA's expected names**
colnames(magma_input) <- c("SNP", "P")



# Write MAGMA-compatible tab-delimited file
write.table(magma_input,
            file = paste0(path3, "BRCA_GWAS_eQTL_PancanQTL_20251201.txt"),
            sep = " ",
            quote = FALSE,
            row.names = FALSE, 
            col.names = TRUE)

```


* why so many rows with `P_value` >1? 



```{r, message=F, warning=F, eval=F}
df <- BRCA_GWAS_eQTL2  

# Extract chromosome and base pair from GWAS_SNP_pos
library(tidyr)
df <- df %>%
  separate(eQTL_SNP_pos, into = c("CHR", "BP"), sep = ":", convert = TRUE)

# Select required columns for MAGMA: SNP, CHR, BP, P
fuma_df <- df[, c( "CHR", "GWAS_SNP",  "P_value")]

# Replace 0 p-values with tiny positive number
fuma_df$P_value[fuma_df$P_value == 0] <- 1e-300


########



# Write MAGMA-compatible tab-delimited file
#write.table(fuma_df,
#            file = paste0("C:\\han\\Projects\\2025_03_breast_cancer\\data_for_FUMA\\", "BRCA_GWAS_eQTL_PancanQTL.txt"),
#            sep = "\t",
#            quote = FALSE,
#            row.names = FALSE,
#            col.names = TRUE)

```



```{r, message=F, warning=F}
df_test <- read.table(paste0(path3, "BRCA_GWAS_eQTL_PancanQTL.txt"), header=TRUE, sep="\t", stringsAsFactors=FALSE)
str(df_test)
```


```{r, message=F, warning=F}
df <- read.delim(paste0(path3, "BRCA_GWAS_eQTL_PancanQTL.txt"), header=TRUE)

df$SNP <- trimws(df$SNP)
df$P <- trimws(df$P)

df <- df[grepl("^[0-9eE.-]+$", df$P), ]   # keep only valid numeric P

write.table(df, file = paste0(path3, "BRCA_GWAS_eQTL_PancanQTL2.txt"),
            sep="\t", row.names=FALSE, quote=FALSE)

#pd_summary=read.table("C:\\han\\Dataset\\20250221_PMID_32152537_hmagma\\6382668\\HMAGMA_Protocol_v1.02\\HMAGMA_Protocol\\PD.summary.stat.txt", header=T)
```



## run MAGMA

```{r, message=F, warn=F, eval=F}
# run MAGMA
magma --bfile C:\\han\\Dataset\\20250221_PMID_32152537_hmagma\\g1000_eur\\g1000_eur --pval C:\\han\\Projects\\2025_03_breast_cancer\\data_for_hmagma\\BRCA_GWAS_eQTL.txt use=tag_SNP, p-value ncol=1092 --gene-annot C:\\han\\Projects\\2025_03_breast_cancer\\data_for_hmagma\\BRSTCancer.transcript.annot --out C:\\han\\Projects\\2025_03_breast_cancer\\data_for_hmagma\\magma_output\\
```
