---
title: "run magma on GWAS summary data only"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
date: "2026-03-30"
---





## Prepare libraries and directory ● Timing 5 min


```{r, message=F, warning=F}
rm(list=ls())
set.seed(123)
options(stringsAsFactors=F)
library(GenomicRanges)
library(biomaRt)
library(dplyr)
```




## Create GenomicRanges (GRanges) objects for exon and promoter coordinates ● Timing ~5 min

```{r, message=F, warning=F}
# 3 Read in exonic and promoter coordinate ﬁles by entering the following commands
path="C:\\han\\Dataset\\20250221_PMID_32152537_hmagma\\6382668\\HMAGMA_Protocol_v1.02\\HMAGMA_Protocol\\"
path2="C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_Only\\"
#path3="C:\\han\\Projects\\2025_03_breast_cancer\\data_for_hmagma\\"
#path4="C:\\han\\Dataset\\20251106_PancanQTL_v2\\BRCA.eQTL-GWAS.xls\\"


exon <- read.table(paste0(path, "Gencode26_exon.bed"), header = FALSE)
dim(exon)
head(exon)
```



```{r, message=F, warning=F}

file.exists(paste0(path, "Gencode26_exon.bed")) # check if the file exists 

exon$V1 <- sub("^", "chr", exon$V1)
promoter <- read.table(paste(path, "Gencode26_promoter.bed", sep=""))
promoter$V1 <- sub("^", "chr", promoter$V1)

# 4 Create a GRanges object for exon and promoter deﬁnitions.
exonranges <- GRanges(exon[,1],IRanges(exon[,2],exon[,3]),gene=exon[,4])
promoterranges <- GRanges(promoter[,1], IRanges(promoter[,2], promoter[,3]), gene=promoter[,4])

# 5 Save exon and promoter GRanges objects as a .rda ﬁle for future use.
#save(exonranges, promoterranges, file=paste(path2, "exon_promoranges.rda", sep=""))

exonranges
promoterranges
```



* "Obtain gene and exon coordinates from Gencode version 26 by using the [following link](https://www.
gencodegenes.org/human/release_26lift37.html). We downloaded the GFT ﬁle for basic gene
annotation and converted them to .bed ﬁles. Promoters were deﬁned as 2 kb upstream to every
transcription start site. Gene, exon and promoter coordinate ﬁles are listed as Gencode26_gene.bed,
Gencode26_exon.bed and Gencode26_promoter.bed, respectively, in the Zenodo repository. Columns
for these ﬁles are deﬁned as [chromosome, start, end, gene]."

**Genome assembly hg19 was used**. 







## Generate a GRanges object for SNP coordinates ● Timing ~5 min

```{r, message=F, warning=F}
#Generate a GRanges object for SNP coordinates ● Timing ~5 min
#6 Read in the SNP annotation ﬁle by using the following command.
snps <- read.table(paste(path, "EUR.bim.gz", sep=""))
snps <- snps[,c(1,2,4)]
colnames(snps) <- c("chr","SNP","Position")
snps$chr <- sub("^", "chr", snps$chr)


#7 Create a GRanges object for SNP annotations.
snps <- GRanges(snps$chr, IRanges(snps$Position, snps$Position), rsid=
snps$SNP)
#8 Save the SNP GRanges object as a .rda ﬁle for future use.
#save(snps,file=paste(path2, "snps.rda", sep=""))
snps
```



## Assign SNPs (in EUR.bim) to genes by overlapping SNPs with exons and promoters ● Timing 10–15 min

```{r, message=F, warning=F}
#9 Identify genes that map to SNPs residing in exons by overlapping the GRanges object for SNPs with
#the GRanges object for exons.
olap <- findOverlaps(snps,exonranges)
snpexon <- snps[queryHits(olap)]
mcols(snpexon) <- cbind(mcols(snpexon), mcols(exonranges[subjectHits(olap)]))
snpexon <- snpexon[seqnames(snpexon)!="chrX"]


#10 Overlap the GRanges object for promoters with the GRanges object for SNPs. Similar to the code
#above, this command will identify genes that map to SNPs residing in promoter regions.
olap <- findOverlaps(snps,promoterranges)
snpro <- snps[queryHits(olap)]
mcols(snpro) <- cbind(mcols(snpro),mcols(promoterranges[subjectHits(olap)]))
snpro <- snpro[seqnames(snpro)!="chrX"]

#11 Save SNPs overlapping with exons and promoters as a .rda ﬁle.
#save(snpro, snpexon, file=paste(path2, "snp_locating_in_exon_promoter_transcript_level.rda", sep=""))
```


* “Find SNPs that fall inside any exon, attach the exon’s annotation (like gene ID) to those SNPs, and remove SNPs on chromosome X.”



```{r, message=F, warning=F}
#21 Integrate SNP-gene relationships derived from exons, promoters and Hi-C interaction data.
load(paste(path2, "snp_locating_in_exon_promoter_transcript_level.rda", sep=""))


snpromat <- unique(data.frame(rsid=snpro$rsid, ensg=snpro$gene))
snpexonmat <- unique(data.frame(rsid=snpexon$rsid, ensg=snpexon$gene))
head(snpexonmat)
#save(snpexonmat, file=paste(path2, "SNP_to_transcript_comb.rda", sep=""))

cat("Total mapped SNPs: ", nrow(snpexonmat), "\n")

```




```{r, message=F, warning=F, eval=F}
#22 Aggregate SNP-gene relationships to generate the variant-gene annotation ﬁle compatible
#with MAGMA.
snpagg <- aggregate(snpexonmat, list(snpexonmat$ensg), unique)


#23 Read in the gene deﬁnition ﬁle.
genedef <- read.table(paste(path, "Gencode26_gene.bed", sep=""))
colnames(genedef) <- c("chr", "start", "end", "ensg")


#24 Create an index column from the gene deﬁnition ﬁle consisting of gene chromosomal location, start
#and end.
genedef <- genedef [grep("chr", genedef$chr),]
genedef$chr <- unlist(lapply(strsplit(genedef$chr, "chr"), '[[', 2))
genedef$index <- paste(genedef$chr, genedef$start, genedef$end, sep=":")


#25 Attach the index column from Step 24 to the variant-gene annotation ﬁle from Step 22.
snpagg$index <- genedef[match(snpagg$ensg, genedef$ensg),"index"]


#26 Remove any missing values from the variant-gene annotation ﬁle.
snpagg <- snpagg[!is.na(snpagg$index),]

#27 Subset gene, gene location and SNPs from the variant-gene annotation ﬁle.
snpannot <- snpagg[,c("ensg", "index", "rsid")]

#28 Save the variant-gene annotation ﬁle in an executable format.
writable <- format(snpannot)
write.table(writable, file=paste(path2, "SNP_aggregate_transcript.txt", sep=""), quote=F, row.names=F, col.names=F, sep="\t")

#system(paste0("sed -e `s/, /\t/g` < SNP_aggregate_transcript",".txt >","Adultbrain.transcript.annot"))
system("sed -e `s/, /\\t/g` < SNP_aggregate_transcript.txt > transcript.annot")


```


```{r, message=F, warning=F, eval=F}
txt <- readLines(paste0(path2, "SNP_aggregate_transcript.txt"))
txt <- gsub(", ", "\t", txt)
#writeLines(txt, paste0(path2,"transcript.annot"))
```


```{r, message=F, warning=F, eval=F}
CGS03_gwas_summary=as_tibble(read.table("C:\\han\\Dataset\\20260326_TCGD\\CGS03.gwas_summary.txt", header=T))
head(CGS03_gwas_summary)
summary(CGS03_gwas_summary$p.value)
sum(is.na(CGS03_gwas_summary))

CGS03_gwas_summary$p.value[CGS03_gwas_summary$p.value == 0]=1e-100
summary(CGS03_gwas_summary$p.value)
#write.table(CGS03_gwas_summary, file="C:\\han\\Dataset\\20260326_TCGD\\CGS03.gwas_summary_for_magma.txt")

CGS03_gwas_summary_for_magma=as_tibble(read.table("C:\\han\\Dataset\\20260326_TCGD\\CGS03.gwas_summary_for_magma.txt", header=T))

# Ensure it's a data.table for clean export
setDT(CGS03_gwas_summary_for_magma)
# 1. Create a minimal table with ONLY the columns MAGMA needs
magma_input <- CGS03_gwas_summary_for_magma[, .(rsid, p.value)]
# 2. Write it to a file WITHOUT quotes and with a simple TAB separator
#fwrite(magma_input, "C:\\han\\Dataset\\20260326_TCGD\\Breast_Cancer_Clean_MAGMA.txt", 
#       sep = "\t", 
#       quote = FALSE, 
#       row.names = FALSE)
```




```{r, message=F, warning=F, eval=F}
############# pan-ukbb
library(data.table)
phenotype_manifest=fread("C:\\han\\Dataset\\pan_ukbb\\phenotype_manifest.tsv")
phenotype_manifest %>% filter(category=="Breast")

phenotype_manifest %>% filter(phenocode=="174.1") %>% select(aws_path)
```






```{r, message=F, warning=F, eval=F}
# run MAGMA
magma --bfile C:\\han\\Dataset\\20250221_PMID_32152537_hmagma\\g1000_eur\\g1000_eur --pval C:\\han\\Dataset\\20260326_TCGD\\Breast_Cancer_Clean_MAGMA.txt use=rsid, p.value N=139274 --gene-annot C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_Only\\transcript.annot --out C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_Only\\magma_output\\

```