---
title: "4/9/2026"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
date: "2026-04-09"
---

## Prepare libraries and directory ● Timing 5 min


```{r, message=F, warning=F}
rm(list=ls())
set.seed(123)
options(stringsAsFactors=F)
library(GenomicRanges)
library(biomaRt)
library(dplyr)
```




## Create GenomicRanges (GRanges) objects for exon and promoter coordinates ● Timing ~5 min

```{r, message=F, warning=F}
# 3 Read in exonic and promoter coordinate ﬁles by entering the following commands
path="C:\\han\\Dataset\\20250221_PMID_32152537_hmagma\\6382668\\HMAGMA_Protocol_v1.02\\HMAGMA_Protocol\\"
path2="C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_HiC_eQTL\\"
path3="C:\\han\\Dataset\\20251106_PancanQTL_v2\\BRCA.eQTL_GWAS\\"

exon <- read.table(paste0(path, "Gencode26_exon.bed"), header = FALSE)
dim(exon)
head(exon)
```



```{r, message=F, warning=F}

file.exists(paste0(path, "Gencode26_exon.bed")) # check if the file exists 

exon$V1 <- sub("^", "chr", exon$V1)
promoter <- read.table(paste(path, "Gencode26_promoter.bed", sep=""))
promoter$V1 <- sub("^", "chr", promoter$V1)

# 4 Create a GRanges object for exon and promoter deﬁnitions.
exonranges <- GRanges(exon[,1],IRanges(exon[,2],exon[,3]),gene=exon[,4])
promoterranges <- GRanges(promoter[,1], IRanges(promoter[,2], promoter[,3]), gene=promoter[,4])

# 5 Save exon and promoter GRanges objects as a .rda ﬁle for future use.
#save(exonranges, promoterranges, file=paste(path2, "exon_promoranges.rda", sep=""))

exonranges
promoterranges
```





## Generate a GRanges object for SNP coordinates ● Timing ~5 min

```{r, message=F, warning=F}
#Generate a GRanges object for SNP coordinates ● Timing ~5 min
#6 Read in the SNP annotation ﬁle by using the following command.
snps <- read.table(paste(path, "EUR.bim.gz", sep=""))
snps <- snps[,c(1,2,4)]
colnames(snps) <- c("chr","SNP","Position")
snps$chr <- sub("^", "chr", snps$chr)

cat("Total number of snps: ", nrow(snps), "\n")
#7 Create a GRanges object for SNP annotations.
snps <- GRanges(snps$chr, IRanges(snps$Position, snps$Position), rsid=
snps$SNP)
#8 Save the SNP GRanges object as a .rda ﬁle for future use.
#save(snps,file=paste(path2, "snps.rda", sep=""))
snps
```



## Assign SNPs (in EUR.bim) to genes by overlapping SNPs with exons and promoters ● Timing 10–15 min

```{r, message=F, warning=F}
#9 Identify genes that map to SNPs residing in exons by overlapping the GRanges object for SNPs with
#the GRanges object for exons.
olap <- findOverlaps(snps,exonranges)
snpexon <- snps[queryHits(olap)]
mcols(snpexon) <- cbind(mcols(snpexon), mcols(exonranges[subjectHits(olap)]))
snpexon <- snpexon[seqnames(snpexon)!="chrX"]


#10 Overlap the GRanges object for promoters with the GRanges object for SNPs. Similar to the code above, this command will identify genes that map to SNPs residing in promoter regions.
olap <- findOverlaps(snps,promoterranges)
snpro <- snps[queryHits(olap)]
mcols(snpro) <- cbind(mcols(snpro),mcols(promoterranges[subjectHits(olap)]))
snpro <- snpro[seqnames(snpro)!="chrX"]



#11 Save SNPs overlapping with exons and promoters as a .rda ﬁle.
#save(snpro, snpexon, file=paste(path2, "snp_locating_in_exon_promoter.rda", sep=""))
```


## Assign unmapped SNPs (in EUR.bim) to genes on the basis of Hi-C interaction data ● Timing
30–45 min

```{r, message=F, warning=F}
# 12 After identiﬁcation of SNPs that reside in exons and/or promoters, there will be a subset of SNPs
# that do not overlap with either (e.g., intergenic or intronic SNPs). In the following steps, we match
# those SNPs to cognate genes on the basis of Hi-C interaction data. Identify unmapped SNPs by
#using the code below.

snpranges <- snps[!(snps$rsid %in% snpexon$rsid),]
snpranges <- snpranges[!(snpranges$rsid %in% snpro$rsid),]

snpranges
#13 Save unmapped SNPs as a .rda ﬁle.
#save(snpranges, file=paste(path2, "non_exonic_promoter_snp.rda", sep=""))

cat( "exon+promoter mapped snps: ", (1-8916368/9822123)*100, "%", "\n")

```



```{r, message=F, warning=F, eval=T}
#14 Read in primary tumor Hi-C data. The ﬁrst three columns represent genomic coordinates [chr, start,
#end] of loop anchor 1, and the last three columns represent genomic coordinates of loop anchor 
path4="C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_HiC\\peaks_in_paper\\"
hic <- read.table(paste0(path4, "GSM8138144_PT1_loops_q0.98_filt.txt"), header = T)


#15 Restructure Hi-C data to account for both anchors 1 and 2 by using the command below.
hic.int1 <- hic[,1:6]
hic.int2 <- hic[,c(4:6,1:3)]
colnames(hic.int1) = colnames(hic.int2) = c("chrom1", "start1", "end1", "chrom2", "start2", "end2")
hic.comb <- rbind(hic.int1, hic.int2)
#hic.comb=hic.comb[-1,]
head(hic.comb)

########### check if  hic.comb has NA or non-numeric values and remove them 
sum(is.na(hic.comb$start1))
sum(is.na(hic.comb$end1))
sum(is.na(hic.comb$start2))
sum(is.na(hic.comb$end2))

bad_rows <- !grepl("^[0-9]+$", hic.comb$start1) | !grepl("^[0-9]+$", hic.comb$end1)
sum(bad_rows)
if (sum(bad_rows)>0)
{
hic.comb[bad_rows, ]
hic.comb=hic.comb[-which(bad_rows==T),]
}
#16 Generate a GRanges object for Hi-C data by using the command below.
hicranges <- GRanges(hic.comb$chrom1, IRanges(as.numeric(hic.comb$start1), as.numeric(hic.comb$end1)), int1=hic.comb$start2,int2=hic.comb$end2)
hicranges
```




```{r, message=F, warning=F}
#17 Identify promoter-anchored interactions by overlapping loop anchor 1 with promoters.
olap <- findOverlaps(hicranges,promoterranges)
generanges <- hicranges[queryHits(olap)]
mcols(generanges) <- cbind(mcols(hicranges[queryHits(olap)]), mcols(promoterranges[subjectHits(olap)]))
generanges
```



```{r, message=F, warning=F}

#18 Reverse the order of the GRanges object from Step 17. The generanges object created in Step 17
#is in a format of loop anchor 1, loop anchor 2, gene name. We reverse the order of loop
#anchors so that the resulting genebed object has a format of loop anchor 2, loop anchor 1,
#gene name.
genebed <- data.frame(chr=seqnames(generanges), snp.start=generanges$int1, snp.end=generanges$int2, gene.start=start(generanges), gene.end=start(generanges)+width(generanges)-1, ensg=generanges$gene)  # anchor 2 information of int 1,2 are used.  

genebed <- unique(genebed)
str(genebed)
head(genebed)

genebed$snp.start <- as.numeric(as.character(genebed$snp.start))
genebed$snp.end   <- as.numeric(as.character(genebed$snp.end))

#19 Create a GRanges object from Step 18.
genesnpranges <- GRanges(genebed$chr, IRanges(genebed$snp.start,genebed$snp.end), ensg=genebed$ensg)  # anchor 2 information is used 
genesnpranges
```



```{r, message=F, warning=F}

#20 Overlap unmapped SNPs from Step 12 with loop anchor 2 from Step 19. This step assigns
#SNPs (located at loop anchor 2) to the genes with which they interact (located at loop
#anchor 1).
olap <- findOverlaps(snpranges,genesnpranges)
snpint <- snpranges[queryHits(olap)]
mcols(snpint) <- cbind(mcols(snpranges[queryHits(olap)]), mcols(genesnpranges[subjectHits(olap)]))
snpint
#save(snpint, file=paste(path2, "PT1_Hi-C_interacting_snp.rda", sep=""))
file.exists(paste0(path2, "PT1_Hi-C_interacting_snp.rda"))
```

## Assign unmapped SNPs (in EUR.bim) to genes based on eQTL


```{r, message=F, warning=F}
#breast_cancer_eqtl=as_tibble(read.csv(paste0(path3, "BRCA.eQTL_GWAS.csv", header=T)))
breast_cancer_eqtl=as_tibble(read.csv("C:\\han\\Dataset\\20251106_PancanQTL_v2\\BRCA.eQTL_GWAS\\BRCA.eQTL_GWAS.csv", header=T))

library(tidyr)
# 1. Clean the eQTL data: Split position and handle column names
# We use eQTL_SNP as rsid and eGene as the gene identifier
eqtl_processed <- breast_cancer_eqtl %>%
  dplyr::select(rsid = eQTL_SNP, pos_raw = eQTL_SNP_pos, gene_symbol = eGene) %>%
  separate(pos_raw, into = c("chr", "pos"), sep = ":") %>%
  mutate(
    chr = paste0("chr", chr), # Add 'chr' prefix to match snpint
    pos = as.integer(pos)
  )

# 2. Map Gene Symbols to ENSG (using an external mapping or  existing data)

 library(org.Hs.eg.db)
 mapping <- select(org.Hs.eg.db, keys=unique(eqtl_processed$gene_symbol), 
                   columns="ENSEMBL", keytype="SYMBOL")
 eqtl_processed <- eqtl_processed %>% 
   left_join(mapping, by=c("gene_symbol"="SYMBOL")) %>%
   mutate(ensg = ENSEMBL) %>% filter(!is.na(ensg))


 
 # Create the GRanges object
eqtl_snpint <- GRanges(
  seqnames = eqtl_processed$chr,
  ranges = IRanges(start = eqtl_processed$pos, end = eqtl_processed$pos),
  rsid = eqtl_processed$rsid,
  ensg = eqtl_processed$ensg 
)

#save(eqtl_snpint, file=paste(path2, "eqtl_snpint.rda", sep=""))
file.exists(paste0(path2, "eqtl_snpint.rda"))
```




```{r, message=F, warning=F}
#21 Integrate SNP-gene relationships derived from exons, promoters and Hi-C interaction data.
load(paste(path2, "PT1_Hi-C_interacting_snp.rda", sep=""))
load(paste(path2, "snp_locating_in_exon_promoter.rda", sep=""))
load(paste(path2, "eqtl_snpint.rda", sep=""))

# 1. Convert Hi-C snpint to dataframe
snpdat <- data.frame(chr=seqnames(snpint),cbp=start(snpint), rsid=snpint$rsid, ensg=snpint$ensg)
head(snpdat)

# 2. Extract Promoter and Exon matrices
snpromat <- unique(data.frame(rsid=snpro$rsid, ensg=snpro$gene))
snpexonmat <- unique(data.frame(rsid=snpexon$rsid, ensg=snpexon$gene))

# 3.  Create the eQTL matrix
snpeqtlmat <- unique(data.frame(rsid = eqtl_snpint$rsid, ensg = eqtl_snpint$ensg))

# 4. Combine ALL sources (Hi-C + Exon + Promoter + eQTL)
snpcomb <- unique(rbind(snpdat[,3:4], snpromat, snpexonmat, snpeqtlmat))
head(snpcomb)
#save(snpcomb, file=paste(path2, "PT1_Hi-C_SNP_comb.rda", sep=""))
cat("Exon+Promoter+Hi-C+eQTL: Total mapped SNPs: ", nrow(snpcomb), "\n")

# 4.  Exon + Promoter 
snpcomb <- unique(rbind(snpromat, snpexonmat))
head(snpcomb)
#save(snpcomb, file=paste(path2, "PT1_Hi-C_SNP_comb.rda", sep=""))
cat("Exon+Promoter: Total mapped SNPs: ", nrow(snpcomb), "\n")


# 4. Exon + Promoter + eQTL
snpcomb <- unique(rbind(snpromat, snpexonmat, snpeqtlmat))
head(snpcomb)
#save(snpcomb, file=paste(path2, "PT1_Hi-C_SNP_comb.rda", sep=""))
cat("Exon+Promoter+eQTL: Total mapped SNPs: ", nrow(snpcomb), "\n")



# 4. Combine ALL sources ( Exon + Promoter + HiC)
snpcomb <- unique(rbind(snpdat[,3:4], snpromat, snpexonmat))
head(snpcomb)
#save(snpcomb, file=paste(path2, "PT1_Hi-C_SNP_comb.rda", sep=""))
cat("Exon+Promoter+Hi-C: Total mapped SNPs: ", nrow(snpcomb), "\n")

mapped_snps=data.frame(conditions=c("Exon+Promoter", "Exon+Promoter+Hi-C", "Exon+Promoter+eQTL", "Exon+Promoter+Hi-C+eQTL"), 
                     num_snps=c(
length(unique(c(snpromat$rsid, snpexonmat$rsid))), 
length(unique(c(snpdat$rsid, snpromat$rsid, snpexonmat$rsid))), 
length(unique(c(snpromat$rsid, snpexonmat$rsid, snpeqtlmat$rsid))), 
length(unique(c(snpdat$rsid, snpromat$rsid, snpexonmat$rsid, snpeqtlmat$rsid)))))

mapped_snps
```





```{r, message=F, warning=F}
ep_snps=unique(c(snpromat$rsid, snpexonmat$rsid))
ep_hic_snps=unique(c(snpdat$rsid, snpromat$rsid, snpexonmat$rsid))
ep_eqtl_snps=unique(c(snpromat$rsid, snpexonmat$rsid, snpeqtlmat$rsid))
ep_hic_eqtl_snps=unique(c(snpdat$rsid, snpromat$rsid, snpexonmat$rsid, snpeqtlmat$rsid))


sum(ep_snps %in% ep_hic_eqtl_snps)==length(ep_snps)
sum(ep_snps %in% snps$rsid)/length(ep_snps)
sum(ep_hic_eqtl_snps %in% snps$rsid)/length(ep_hic_eqtl_snps)
```






```{r, message=F, warning=F, eval=T}
############ Exome+Promoter #############
load(paste(path2, "snp_locating_in_exon_promoter.rda", sep=""))

# 2. Extract Promoter and Exon matrices
snpromat <- unique(data.frame(rsid=snpro$rsid, ensg=snpro$gene))
snpexonmat <- unique(data.frame(rsid=snpexon$rsid, ensg=snpexon$gene))

# 4.  Exon + Promoter 
snpcomb <- unique(rbind(snpromat, snpexonmat))
head(snpcomb)
sum(snpcomb$rsid %in% snps$rsid)/nrow(snpcomb) # how many snps are in reference panel 


#22 Aggregate SNP-gene relationships to generate the variant-gene annotation ﬁle compatible
#with MAGMA.
snpagg <- aggregate(snpcomb, list(snpcomb$ensg), unique)

#23 Read in the gene deﬁnition ﬁle.
genedef <- read.table(paste(path, "Gencode26_gene.bed", sep=""))
colnames(genedef) <- c("chr", "start", "end", "ensg")


#24 Create an index column from the gene deﬁnition ﬁle consisting of gene chromosomal location, start
#and end.
genedef <- genedef [grep("chr", genedef$chr),]
genedef$chr <- unlist(lapply(strsplit(genedef$chr, "chr"), '[[', 2))
genedef$index <- paste(genedef$chr, genedef$start, genedef$end, sep=":")


#25 Attach the index column from Step 24 to the variant-gene annotation ﬁle from Step 22.
snpagg$index <- genedef[match(snpagg$ensg, genedef$ensg),"index"]


#26 Remove any missing values from the variant-gene annotation ﬁle.
snpagg <- snpagg[!is.na(snpagg$index),]

#27 Subset gene, gene location and SNPs from the variant-gene annotation ﬁle.
snpannot <- snpagg[,c("ensg", "index", "rsid")]

#28 Save the variant-gene annotation ﬁle in an executable format.
writable <- format(snpannot)

#write.table(writable, file=paste(path2, "exome_promoter.txt", sep=""), quote=F, row.names=F, col.names=F, sep="\t")

txt <- readLines(paste0(path2, "exome_promoter.txt"))

txt <- gsub(", ", "\t", txt)


missed_genes <- c(
  "ENSG00000200344", "ENSG00000243749", "ENSG00000272226", "ENSG00000234497",
  "ENSG00000212601", "ENSG00000206832", "ENSG00000236887", "ENSG00000224690",
  "ENSG00000232536", "ENSG00000271748", "ENSG00000232519", "ENSG00000228538",
  "ENSG00000230274", "ENSG00000263651", "ENSG00000251804", "ENSG00000264146",
  "ENSG00000251356", "ENSG00000199944", "ENSG00000086232", "ENSG00000213700",
  "ENSG00000266344", "ENSG00000253068", "ENSG00000254732", "ENSG00000162231",
  "ENSG00000252709", "ENSG00000254990", "ENSG00000254450", "ENSG00000118137",
  "ENSG00000196531", "ENSG00000257604", "ENSG00000221560", "ENSG00000102468",
  "ENSG00000102531", "ENSG00000242135", "ENSG00000080824", "ENSG00000270108",
  "ENSG00000260274", "ENSG00000182511", "ENSG00000260494", "ENSG00000175938",
  "ENSG00000183011", "ENSG00000266356", "ENSG00000267340", "ENSG00000265298",
  "ENSG00000182687", "ENSG00000267783", "ENSG00000269181", "ENSG00000268457",
  "ENSG00000221039"
)
ep_missed_genes=snpannot %>% filter(ensg %in% missed_genes)


#writeLines(txt, paste0(path2,"exome_promoter.annot"))

################ run magma ##############
#C:\han\softwares\magma_v1.10_win>magma --bfile C:\\han\\Dataset\\20250221_PMID_32152537_hmagma\\g1000_eur\\g1000_eur --pval C:\\han\\Dataset\\20260326_TCGD\\Breast_Cancer_Clean_MAGMA.txt use=rsid, p.value N=139274 --gene-annot C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_HiC_eQTL\\exome_promoter.annot --out C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_HiC_eQTL\\magma_output_exome_promoter\\

```



```{r, message=F, warning=F, eval=F}
############ Exome+Promoter+Hi-C #############
load(paste(path2, "snp_locating_in_exon_promoter.rda", sep=""))
load(paste(path2, "PT1_Hi-C_interacting_snp.rda", sep=""))

# 1. Convert Hi-C snpint to dataframe
snpdat <- data.frame(chr=seqnames(snpint),cbp=start(snpint), rsid=snpint$rsid, ensg=snpint$ensg)
head(snpdat)

# 2. Extract Promoter and Exon matrices
snpromat <- unique(data.frame(rsid=snpro$rsid, ensg=snpro$gene))
snpexonmat <- unique(data.frame(rsid=snpexon$rsid, ensg=snpexon$gene))

# 4.  Exon + Promoter+HiC 
snpcomb <- unique(rbind(snpdat[,3:4], snpromat, snpexonmat))
head(snpcomb)


#22 Aggregate SNP-gene relationships to generate the variant-gene annotation ﬁle compatible
#with MAGMA.
snpagg <- aggregate(snpcomb, list(snpcomb$ensg), unique)

#23 Read in the gene deﬁnition ﬁle.
genedef <- read.table(paste(path, "Gencode26_gene.bed", sep=""))
colnames(genedef) <- c("chr", "start", "end", "ensg")


#24 Create an index column from the gene deﬁnition ﬁle consisting of gene chromosomal location, start
#and end.
genedef <- genedef [grep("chr", genedef$chr),]
genedef$chr <- unlist(lapply(strsplit(genedef$chr, "chr"), '[[', 2))
genedef$index <- paste(genedef$chr, genedef$start, genedef$end, sep=":")


#25 Attach the index column from Step 24 to the variant-gene annotation ﬁle from Step 22.
snpagg$index <- genedef[match(snpagg$ensg, genedef$ensg),"index"]


#26 Remove any missing values from the variant-gene annotation ﬁle.
snpagg <- snpagg[!is.na(snpagg$index),]

#27 Subset gene, gene location and SNPs from the variant-gene annotation ﬁle.
snpannot <- snpagg[,c("ensg", "index", "rsid")]

#28 Save the variant-gene annotation ﬁle in an executable format.
writable <- format(snpannot)

write.table(writable, file=paste(path2, "exome_promoter_HiC.txt", sep=""), quote=F, row.names=F, col.names=F, sep="\t")

txt <- readLines(paste0(path2, "exome_promoter_HiC.txt"))

txt <- gsub(", ", "\t", txt)

writeLines(txt, paste0(path2,"exome_promoter_HiC.annot"))

################ run magma ##############
#C:\han\softwares\magma_v1.10_win>magma --bfile C:\\han\\Dataset\\20250221_PMID_32152537_hmagma\\g1000_eur\\g1000_eur --pval C:\\han\\Dataset\\20260326_TCGD\\Breast_Cancer_Clean_MAGMA.txt use=rsid, p.value N=139274 --gene-annot C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_HiC_eQTL\\exome_promoter.annot --out C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_HiC_eQTL\\magma_output_exome_promoter\\

```



```{r, message=F, warning=F, eval=F}
############ Exome+Promoter+eQTL #############
load(paste(path2, "snp_locating_in_exon_promoter.rda", sep=""))
load(paste(path2, "eqtl_snpint.rda", sep=""))


# 2. Extract Promoter and Exon matrices
snpromat <- unique(data.frame(rsid=snpro$rsid, ensg=snpro$gene))
snpexonmat <- unique(data.frame(rsid=snpexon$rsid, ensg=snpexon$gene))

# 3.  Create the eQTL matrix
snpeqtlmat <- unique(data.frame(rsid = eqtl_snpint$rsid, ensg = eqtl_snpint$ensg))


# 4.  Exon + Promoter+eQTL 
snpcomb <- unique(rbind(snpromat, snpexonmat, snpeqtlmat))
head(snpcomb)


#22 Aggregate SNP-gene relationships to generate the variant-gene annotation ﬁle compatible
#with MAGMA.
snpagg <- aggregate(snpcomb, list(snpcomb$ensg), unique)

#23 Read in the gene deﬁnition ﬁle.
genedef <- read.table(paste(path, "Gencode26_gene.bed", sep=""))
colnames(genedef) <- c("chr", "start", "end", "ensg")


#24 Create an index column from the gene deﬁnition ﬁle consisting of gene chromosomal location, start
#and end.
genedef <- genedef [grep("chr", genedef$chr),]
genedef$chr <- unlist(lapply(strsplit(genedef$chr, "chr"), '[[', 2))
genedef$index <- paste(genedef$chr, genedef$start, genedef$end, sep=":")


#25 Attach the index column from Step 24 to the variant-gene annotation ﬁle from Step 22.
snpagg$index <- genedef[match(snpagg$ensg, genedef$ensg),"index"]


#26 Remove any missing values from the variant-gene annotation ﬁle.
snpagg <- snpagg[!is.na(snpagg$index),]

#27 Subset gene, gene location and SNPs from the variant-gene annotation ﬁle.
snpannot <- snpagg[,c("ensg", "index", "rsid")]

#28 Save the variant-gene annotation ﬁle in an executable format.
writable <- format(snpannot)

write.table(writable, file=paste(path2, "exome_promoter_eQTL.txt", sep=""), quote=F, row.names=F, col.names=F, sep="\t")

txt <- readLines(paste0(path2, "exome_promoter_eQTL.txt"))

txt <- gsub(", ", "\t", txt)

writeLines(txt, paste0(path2,"exome_promoter_eQTL.annot"))

################ run magma ##############
#C:\han\softwares\magma_v1.10_win>magma --bfile C:\\han\\Dataset\\20250221_PMID_32152537_hmagma\\g1000_eur\\g1000_eur --pval C:\\han\\Dataset\\20260326_TCGD\\Breast_Cancer_Clean_MAGMA.txt use=rsid, p.value N=139274 --gene-annot C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_HiC_eQTL\\exome_promoter.annot --out C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_HiC_eQTL\\magma_output_exome_promoter\\

```




```{r, message=F, warning=F, eval=T}
############ Exome+Promoter+HiC+eQTL #############
load(paste(path2, "snp_locating_in_exon_promoter.rda", sep=""))
load(paste(path2, "PT1_Hi-C_interacting_snp.rda", sep=""))
load(paste(path2, "eqtl_snpint.rda", sep=""))


# 1. Convert Hi-C snpint to dataframe
snpdat <- data.frame(chr=seqnames(snpint),cbp=start(snpint), rsid=snpint$rsid, ensg=snpint$ensg)
head(snpdat)

# 2. Extract Promoter and Exon matrices
snpromat <- unique(data.frame(rsid=snpro$rsid, ensg=snpro$gene))
snpexonmat <- unique(data.frame(rsid=snpexon$rsid, ensg=snpexon$gene))

# 3.  Create the eQTL matrix
snpeqtlmat <- unique(data.frame(rsid = eqtl_snpint$rsid, ensg = eqtl_snpint$ensg))


# 4.  Exon + Promoter+HiC+eQTL 
snpcomb <- unique(rbind(snpdat[,3:4], snpromat, snpexonmat, snpeqtlmat))
head(snpcomb)


## check one specific gene:  ENSG00000086232




#22 Aggregate SNP-gene relationships to generate the variant-gene annotation ﬁle compatible
#with MAGMA.
snpagg <- aggregate(snpcomb, list(snpcomb$ensg), unique)

#23 Read in the gene deﬁnition ﬁle.
genedef <- read.table(paste(path, "Gencode26_gene.bed", sep=""))
colnames(genedef) <- c("chr", "start", "end", "ensg")


#24 Create an index column from the gene deﬁnition ﬁle consisting of gene chromosomal location, start
#and end.
genedef <- genedef [grep("chr", genedef$chr),]
genedef$chr <- unlist(lapply(strsplit(genedef$chr, "chr"), '[[', 2))
genedef$index <- paste(genedef$chr, genedef$start, genedef$end, sep=":")


#25 Attach the index column from Step 24 to the variant-gene annotation ﬁle from Step 22.
snpagg$index <- genedef[match(snpagg$ensg, genedef$ensg),"index"]


#26 Remove any missing values from the variant-gene annotation ﬁle.
snpagg <- snpagg[!is.na(snpagg$index),]

#27 Subset gene, gene location and SNPs from the variant-gene annotation ﬁle.
snpannot <- snpagg[,c("ensg", "index", "rsid")]

#28 Save the variant-gene annotation ﬁle in an executable format.
writable <- format(snpannot)

write.table(writable, file=paste(path2, "exome_promoter_HiC_eQTL.txt", sep=""), quote=F, row.names=F, col.names=F, sep="\t")

txt <- readLines(paste0(path2, "exome_promoter_HiC_eQTL.txt"))

txt <- gsub(", ", "\t", txt)

missed_genes <- c(
  "ENSG00000200344", "ENSG00000243749", "ENSG00000272226", "ENSG00000234497",
  "ENSG00000212601", "ENSG00000206832", "ENSG00000236887", "ENSG00000224690",
  "ENSG00000232536", "ENSG00000271748", "ENSG00000232519", "ENSG00000228538",
  "ENSG00000230274", "ENSG00000263651", "ENSG00000251804", "ENSG00000264146",
  "ENSG00000251356", "ENSG00000199944", "ENSG00000086232", "ENSG00000213700",
  "ENSG00000266344", "ENSG00000253068", "ENSG00000254732", "ENSG00000162231",
  "ENSG00000252709", "ENSG00000254990", "ENSG00000254450", "ENSG00000118137",
  "ENSG00000196531", "ENSG00000257604", "ENSG00000221560", "ENSG00000102468",
  "ENSG00000102531", "ENSG00000242135", "ENSG00000080824", "ENSG00000270108",
  "ENSG00000260274", "ENSG00000182511", "ENSG00000260494", "ENSG00000175938",
  "ENSG00000183011", "ENSG00000266356", "ENSG00000267340", "ENSG00000265298",
  "ENSG00000182687", "ENSG00000267783", "ENSG00000269181", "ENSG00000268457",
  "ENSG00000221039"
)
ep_hic_eqtl_missed_genes=snpannot %>% filter(ensg %in% missed_genes)


writeLines(txt, paste0(path2,"exome_promoter_HiC_eQTL.annot"))

################ run magma ##############
#C:\han\softwares\magma_v1.10_win>magma --bfile C:\\han\\Dataset\\20250221_PMID_32152537_hmagma\\g1000_eur\\g1000_eur --pval C:\\han\\Dataset\\20260326_TCGD\\Breast_Cancer_Clean_MAGMA.txt use=rsid, p.value N=139274 --gene-annot C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_HiC_eQTL\\exome_promoter.annot --out C:\\han\\Projects\\2025_03_breast_cancer\\Run_MAGMA_on_GWAS_HiC_eQTL\\magma_output_exome_promoter\\

```



```{r, message=F, warning=F}
target_gene="ENSG00000086232"
ep_missed_genes %>% filter(ensg==target_gene)
ep_missed_genes_snps=ep_missed_genes %>% filter(ensg == target_gene) %>% pull(rsid)
sum(ep_missed_genes_snps %in% snps$rsid) ############ is looks all snps are in reference panel, why here it's 0?????????



num_snps_in_reference=numeric()
for (i in 1:length(missed_genes))
{
  ep_missed_genes_snps=ep_missed_genes %>% filter(ensg == missed_genes[i]) %>% pull(rsid)
  num_snps_in_reference[i]=sum(ep_missed_genes_snps %in% snps$rsid)
}



ep_hic_eqtl_missed_genes_snps=ep_hic_eqtl_missed_genes %>% filter(ensg==target_gene)%>% pull(rsid)

num_snps_in_reference2=numeric()
for (i in 1:length(missed_genes))
{
  ep_hic_eqtl_missed_genes_snps=ep_hic_eqtl_missed_genes %>% filter(ensg == missed_genes[i]) %>% pull(rsid)
  num_snps_in_reference2[i]=sum(ep_hic_eqtl_missed_genes_snps %in% snps$rsid)
}

intersect(ep_hic_eqtl_missed_genes_snps[[1]], ep_missed_genes_snps[[1]])
setdiff(ep_hic_eqtl_missed_genes_snps[[1]], ep_missed_genes_snps[[1]])
sum(ep_hic_eqtl_missed_genes_snps %in% snps$rsid)

```
* all snps are found in 